---
title:  "A mathematical model to study defensive metrics in football: individual, collective and game pressures"
subtitle: "SIMULATOR for the Dynamical System of Pressures."
author: 
  name: "Cesar Catalan"
  orcid:  0000-0002-6602-174X 
  email: cecapa2@doctor.upv.es
  affiliations: "IUMPA. Instituto Universitario de Matemática Pura y Aplicada"
keywords:  football, soccer, modelling, pressure, dynamical system
abstract: We present a complete model for the calculation of defensive dynamic pressures that is able to evaluate and analyse the movement of defensive pressures from a spatio-temporal point of view in football, associating it with the equivalence to a Dynamic System of Variable Mass Particles.
version: 1.0
date: July 2025
format:
  html:
    toc: true
    toc-depth: 3
    toc-expand: 3
    toc-title: Contents
    number-sections: true
    code-line-numbers: true #numera las líneas de código
    theme: materia #cosmo #sandstone #journal #litera
    html-math-method: katex #para los entornos matematicos ver que render utilizar
     #css: journal.css (por si queremos añadir uno propio)
    df-print: kable #Para presentar tablas con dataframes
    code-fold: false #Para visualizar/escondeer el código inicialmente
    code-summary: "Show the code"
    tbl-cap-location: top #Ubicacion titulo tablas
    smooth-scroll: true # Enable smooth scrolling within the page
    code-overflow: scroll #wrap Scroll code blocks that exceed width or Wrap lines
execute:
  echo: true       # Muestra el código fuente en el documento final.
  output: true     #  Muestra el resultado de la ejecución del código en el documento final.
  error: true      # Muestra los mensajes de error generados por la ejecución del código.
  warning: false    # No muestra los mensajes de advertencia generados por la ejecución del código.
  #fig-width: ancho  # Ajusta la anchura de las figuras generadas por los bloques de código a ancho.
  #fig-height: alto  # Ajusta la altura de las figuras generadas por los bloques de código a alto.
  cache: false       # Guarda en caché los resultados de la ejecución de los bloques de código.
  enable: false     # Deshabilita la ejecución de los bloques de código.
  include: true    #ejecuta los include que hemos ido añadiendo 
editor: visual
editor_options: 
  chunk_output_type: console
---

# PACKAGES

```{r}
#| format:
#|  html: 
#|    code-fold: true
#|    code-summary: "Show the code"
#|    code-tools: true

# R version: R-4.4.2 (OK)
# R version: R-4.5.1 (Necesary for ggstream package)
# Necesary compatible Rtool (4.5.6608) with R-4.5.1. Version 4.4.6459 doesn't wotk 
# RStudio version: 2024.12.0 Build 467
# Quarto version: 

  ########## PAQUETES EXCLUSIVOS PARA EL SIMULADOR ###########
                    ##REVISAR##
  # instalamos el paquete ggsoccer para dibujar campos de fútbol y sus jugadores
remotes::install_github("torvaney/ggsoccer")
install.packages("ggsoccer")
library(ggsoccer)
if(!require(gganimate))install.packages("gganimate");library("gganimate")
if(!require(av))install.packages("av");library("av")
  # El siguiente paquete es necesario para utilizar las transformaciones de poligonos en el animate y ggforce
devtools::install_github('thomasp85/transformr')  # https://github.com/thomasp85/transformr
  # Este package es para crear formatos GIF
if(!require(gifski))install.packages("gifski");library("gifski")

                   ##REVISAR##

  ########## PAQUETES UTILIZADOS EN KPIs.qmd ################

if(!require(here))install.packages("here");library("here")
if(!require(psych))install.packages("psych");library("psych")
if(!require(tidyverse))install.packages("tidyverse");library("tidyverse")
if(!require(reshape2))install.packages("reshape2");library("reshape2")
if(!require(readr))install.packages("readr");library("readr") #para usar la funci?n read.file()
if(!require(readtext))install.packages("readtext");library("readtext")  #para poder leer el archivo
if(!require(XML))install.packages("XML");library("XML") #para leer archivos .xml y ejecutar FCrSTATS
if(!require(xml2))install.packages("xml2");library("xml2")  #para leer archivos .xml
if(!require(dplyr))install.packages("dplyr");library("dplyr") #para ordenar archivos y manipular datos
if(!require(purrr))install.packages("purrr");library("purrr") #para manipular datos
if(!require(e1071))install.packages("e1071");library("e1071") #para calcular Kurtosis, Fisher, etc.
if(!require(stats))install.packages("stats");library("stats") #para calculo de regresiones lineales
if(!require(Metrics))install.packages("Metrics");library("Metrics") #para calculo de RMSE
if(!require(ggtern))install.packages("ggtern");library("ggtern")  #para el calculo de funciones de densidad weighted NO PARECE FUNCIONAR!!!!
#para graficos
if(!require(ggplot2))install.packages("ggplot2");library("ggplot2") #para representar graficos de varias variables
if(!require(ggpubr))install.packages("ggpubr");library("ggpubr")  #para combinar varios ggplots
if(!require(ggstream))install.packages("ggstream");library("ggstream")  #para graficos tipo stream
if(!require(ggExtra))install.packages("ggExtra");library("ggExtra")  #para graficos tipo marginal
if(!require(gridExtra))install.packages("gridExtra");library("gridExtra") #for adding special plots to ggplot2.
if(!require(scales))install.packages("scales");library("scales")  #para modificar las escalas de los ejes en ggplot2
if(!require(raster))install.packages("raster");library("raster")  #to create maps
if(!require(plotly))install.packages("plotly");library("plotly")  #?3D plots surface. "plot_ly()"
if(!require(listviewer))install.packages("listviewer");library("listviewer") #para ver los typos de trazas en plotly
# schema(jsonedit = interactive()) esta instruccion se usa para mostrar las propiedades de plotly
if(!require(viridis))install.packages("viridis");library("viridis") #color scales
if(!require(RColorBrewer))install.packages("RColorBrewer");library("RColorBrewer")  #color package
# scale_fill_brewer() for box plot, bar plot, violin plot, dot plot, etc
# scale_color_brewer() for lines and points
if(!require(hrbrthemes))install.packages("hrbrthemes");library("hrbrthemes")  #themas
if(!require(ggthemes))install.packages("ggthemes");library("ggthemes")  #para variar los temas de los gráficos
if(!require(ggforce))install.packages("ggforce");library("ggforce") #para dibujar sobre ggplot (voronoi, circle, etc..)
if(!require(latex2exp))install.packages("latex2exp");library("latex2exp") #para utilizar latex en gráficos y expresiones
if(!require(fmsb))install.packages("fmsb");library("fmsb")  #para graficos de radar y spider
if(!require(ggrepel))install.packages("ggrepel");library("ggrepel") #para repeler texto en gráficos
if(!require(raster))install.packages("raster");library("raster")  #to create maps
if(!require(grid))install.packages("grid");library("grid")  # para dibujar

if(!require(jsonlite))install.packages("jsonlite");library("jsonlite")  # para trabajr con archivos .json
if(!require(xml2))install.packages("xml2");library("xml2")  # para trabajr con archivos .json
if(!require(rmarkdown))install.packages("rmarkdown");library("rmarkdown") # para trabajar con Quarto
if(!require(knitr))install.packages("knitr");library("knitr")  # para trabajar con tabas en html
#if(!require(kableExtra))install.packages("kableExtra");library("kableExtra")  # para trabajar con tabas en html

if(!require(chron))install.packages("chron");library("chron")  # para trabajar con variables cronológicas de tiempo

#if(!require(SparkR))install.packages("SparkR");library("SparkR")  # para calcular el acos()
# if(!require(igraph))install.packages("igraph");library("igraph")  # para grafos y network diagrams

```

# OBTAINING DATA OF THE DYNAMIC MODEL

Obtenemos los dataframes para la simulación. Este código solo lo debemos ejecutar en caso de no haber ejecutado el archivo principal KPIs.qmd

```{r}
#| label: Global variables and dataframes
#| eval: false
#| code-fold: true
#| include: false

  #Variables de path ya definidas en KPIs.qmd
pathDataOrig <- "Sample_Game_3/"
pathimages <- "KPIsImages/"
pathvideos <- "KPIsVideos/"
pathdata  <- "KPIsData/"

  #Variables globales ya definidas en KPIs.qmd

    ####################################################################
    # Declaracion de las variables GLOBALES r, h, k, Amin, Amax, medidas del campo, etc.
    ###################################################################
  # Frame frequency: 25 frames/s
h<-25 
  #radius for the dinamical model: 5 meters
r<-5
  #Coef. of influence area
k= pi/3+sqrt(3)/2
  #Miximal influence area
Amin= (pi-k)*r^2
  #Maximal influence area
Amax=pi*r^2
  #Ratio Amin/Amax
Rate <- Amin/Amax
  #Equivalent radious for Pab. pi*req^2=Pab -> req=(Pab/pi)^1/2

# Load dataframes

#METADATA
Meta <- readRDS(str_c(pathdata,"Meta.rds"))

  #PLAYERS
PLAYER <- readRDS(str_c(pathdata,"PLAYER.rds"))

  #EVENTS
EVENT <- readRDS(str_c(pathdata,"EVENT.rds"))
EVENT_lost <- readRDS(str_c(pathdata,"EVENT_Lost.rds"))
EVENT_lost2 <- readRDS(str_c(pathdata,"EVENT_Lost2.rds"))
EVENT_BallLost <- readRDS(str_c(pathdata,"EVENT_BallLost.rds"))
EVENT_CornerKick <- readRDS(str_c(pathdata,"EVENT_CornerKick.rds"))
EVENT_throwIn <- readRDS(str_c(pathdata,"EVENT_throwIn.rds"))

  #FRAMES
nFr_Players_a <- readRDS(str_c(pathdata,"nFr_Players_a.rds"))
nFr_Players_b <- readRDS(str_c(pathdata,"nFr_Players_b.rds"))
nFr_Playersab <- readRDS(str_c(pathdata,"nFr_Players_ab.rds"))

  #KINEMATIC
KINEM <- readRDS(str_c(pathdata,"KINEM.rds"))

  #DYNAMICS
DYNAM <- readRDS(str_c(pathdata,"DYNAM.rds"))
DYNAM_a <- readRDS(str_c(pathdata,"DYNAM_a.rds"))
DYNAM_b <- readRDS(str_c(pathdata,"DYNAM_b.rds"))
DYNAM_ab <- readRDS(str_c(pathdata,"DYNAM_ab.rds"))
DYNAM_AB <- readRDS(str_c(pathdata,"DYNAM_AB.rds"))
DYNAM_BA <- readRDS(str_c(pathdata,"DYNAM_BA.rds"))
DYNAM_CP_AB <- readRDS(str_c(pathdata,"DYNAM_CP_AB.rds"))

  # pitch dimensions 
fPitchXSizeMeters <-Meta %>% filter(Description=="WidthFieldsize") %>% select(Value)
fPitchYSizeMeters <-Meta %>% filter(Description=="HeigthFieldsize") %>% select(Value)
fPitchXSizeMeters <- round(as.numeric(fPitchXSizeMeters),6)
fPitchYSizeMeters <- round(as.numeric(fPitchYSizeMeters),6)
  # mean frames
Fr_1I<-361  # Start of the 1st period
Fr_1F<-69594 # End of the 1st period 
Fr_2I<-70090 # Start of the 2nd period
Fr_2F<-143553 # End of the 2nd period
Fr_I_Act<-Fr_1I   #Frame Inicio seleccionado para el calculo
Fr_F_Act<-Fr_2F  #Fr_2F   #Frame Final seleccionado para el calculo
Fr_Total<-as.integer(Fr_F_Act)-as.integer(Fr_I_Act)+1  #Frame usados para el calculo

```

## Selecting actions and frames

```{r}
#| label: Actions and frames
#| echo: true
#| code-fold: true

# (A1) counter-attack
E <- EVENT_lost2 %>% filter(Id==31)
K <- KINEM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)
D <- DYNAM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)


# (A2) counter-attack
E <- EVENT_lost2 %>% filter(Id==86)
K <- KINEM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)
D <- DYNAM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)

# (A3) counter-attack
E <- EVENT_lost2 %>% filter(Id==92)
K <- KINEM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)
D <- DYNAM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)

# (B1) corner-kick
E <- EVENT_CornerKick %>% filter(id==1)
K <- KINEM %>% filter (Fr>=E$Fr[1] & (Fr<E$Fr[1]+25*5))
D <- DYNAM %>% filter (Fr>=E$Fr[1] & (Fr<E$Fr[1]+25*5)) %>% mutate(id=1)

# (C1) throw-in
E <- EVENT_throwIn %>% filter(id==1)
K <- KINEM %>% filter (Fr>=E$Fr[1] & (Fr<E$Fr[1]+25*5))
D <- DYNAM %>% filter (Fr>=E$Fr[1] & (Fr<E$Fr[1]+25*5)) %>% mutate(id=1)

```

## Preparing pitch

We use the **ggsoccer**\[\^6\] package which produce a pitch with apropiated dimensions and coordinates. \[\^6\]: <https://cran.r-project.org/web/packages/ggsoccer/ggsoccer.pdf>

```{r}
#| label: Preparing the pitch
#| echo: true
#| code-fold: true

  #We can use the "ggsoccer" package
# Pitch <- ggsoccer::annotate_pitch(
#                   colour = "dimgray",#"White",
#                   fill   = "white", #springgreen4", #rgb(0,0.6,0,0.2)
#                   limits = FALSE,
#                     # Tracab reference (origin in the center of the pitch)
#                   dimensions = make_pitch_tracab(fPitchXSizeMeters, fPitchYSizeMeters),
#                   #dimensions=rescale_coordinates(pitch_tracab,pitch_international),
#                   goals=goals_line, #goals_box
#                   linewidth = 0.5,
#                   alpha=0.1,
#                   linetype="solid"
#                 )
  #but weusea list with the dimensions of the Sample3 pitch, changing the origin coordinates fot plot it
Pitch_custom <- list( 
  length = fPitchXSizeMeters,
  width = fPitchYSizeMeters, 
  penalty_box_length = 16.5, 
  penalty_box_width = 40.3, 
  six_yard_box_length = 5.5, 
  six_yard_box_width = 18.3, 
  penalty_spot_distance = 11, 
  goal_width = 7.32, 
  origin_x = -fPitchXSizeMeters/2, 
  origin_y = -fPitchYSizeMeters/2 
  )

# Create a main transparent theme object 
transparent_theme <- theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  axis.text.x = element_blank(), 
  axis.text.y = element_blank(),
  axis.ticks = element_blank(),
  panel.grid = element_blank(),
  axis.line = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA)
 )
```

```{r}
#| label: Defining colours
#| echo: true
#| code-fold: true

  ################################################################################
  # selecting colours for variables and auziliar objects
  ################################################################################

ai_colour <- rgb(0,0,1,alpha=0.75,names="ai color") # para los jugadores a_i
bk_colour <- rgb(150/255,100/255,0,alpha=0.75,names="bk color") # para los jugadores b_k
cj_colour <- rgb(250/255,0,0,alpha=0.75,names="cj color") # para los jugadores c_j
d_colour<- "gray" #color de las lineas que unen los COP
b_colour<- "gray" #color de las bolas de los jugadores con presión
CP_colour<- "darkgreen"  #color de las leyendas de los COPs de los jugadores con presión
                          #color de los bordes de las áreas de presion generadas
p_colour<-rgb(100/255,150/255,0,alpha=255/255,names="p color") #color de las áres de presiones generadas era 64/255 en realidad
v_colour <- "darkblue" #color de las velocidades
a_colour <- "cyan3"  #color de las aceleraciones
L_colour <- rgb(255/255,0,0,alpha=1,names="J color")  #color de Jab
F_colour <- rgb(250/255,200/255,0,alpha=1,names="F color")  #color de Fab
W_colour <- rgb(0,200/255,1,alpha=1,names="W color")  #color de Wab


  ################################################################################
  # selecting linear scales for displaying variables P,J,F y W 
  ################################################################################
# definicion de las escalas para representar P,J,F y W 
S_Max <- 3*r  #asignamos el valor máximo de escala para las variables

s_v <-  S_Max/8       #escala máxima a representar / valor maximo de la variable  
s_a <-  S_Max/8       #aceleraciones
s_P <-  Amax          #Presiones
s_J <-  S_Max/4       #J
s_F <-  S_Max/4     #F
s_W <-  S_Max/4     #W
s_E <-  S_Max/4     #E

```

# BUILDING A DYNAMICAL MODEL

## Didactical.1 frame

Aqui vamos dibujando y creando el modelo dinámico apartir de los jugadores y su cinemática

```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph101-WbA+WBa_Headmap

  # (A1) counter-attack - FIRST Frame
# fr=31
# E <- EVENT_lost2 %>% filter(Id==fr)
# K <- KINEM %>% filter(Fr==E$Fr)
# D <- DYNAM %>% filter(Fr==E$Fr)
  # Selecting the frame to drawing
fr=92500 # (92500 frame of event Id=86)
K <- KINEM %>% filter(Fr==fr)
D <- DYNAM %>% filter(Fr==fr)
DAB <- DYNAM %>% filter(Fr==fr) %>% group_by(Bowing) %>% 
  summarise(pAB=sum(pab), LAB=sum(Lab,na.rm=TRUE), WAB=sum(Wab,na.rm=TRUE)) %>% ungroup()
D_CP_AB <- DYNAM_CP_AB %>% filter(Fr==fr)

# (A1) counter-attack - ALL Frames
# E <- EVENT_lost2 %>% filter(Id==81)
# K <- KINEM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)
# D <- DYNAM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)

Play1 <- ggplot(K) +
    # 1-Drawing the pitch
  #annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing players and ball positions
  geom_point(aes(x=r_x,y=r_y,color=team,size = team),
             alpha=0.5, show.legend = FALSE) +
    # 3-Drawing player numbers
  geom_text(data=K,aes(x = r_x, y = r_y,color= team),
                   label=K$player_shirt, max.overlaps=10, size=3, show.legend = FALSE)+ 
    # 4- (v) Drawing velocities of players as vectors
  geom_segment(data=K,aes(x = r_x, y = r_y, xend=r_x+v_x*s_v, yend=r_y+v_y*s_v),
              color=v_colour,
              arrow=arrow(length = unit(0.2,"cm")),
              lineend="round", linejoin = "round",
              linetype =2,
              linewidth=0.50)+
    # 5- (a) Drawing acelerations of players as vectors
  geom_segment(data=K,aes(x = r_x, y = r_y, xend=r_x+a_x*s_a, yend=r_y+a_y*s_a),
              color=a_colour,
              arrow=arrow(length = unit(0.1,"cm")),
              lineend="round", linejoin = "round",
              linetype =2,
              linewidth=0.50)+
    # 6- Drawing the circumferences (radi=r) for all players
  geom_circle(data=filter(K,K$team!="C"),aes(x0=r_x,y0=r_y,r=r),
              n=25,color="gray",alpha=0.05,linetype=2)+
    # 7- Drawing line between a and b players to represente the CP
  geom_segment(data=D,aes(x = ra_x, y = ra_y, xend=rb_x, yend=rb_y),
             color= d_colour,linetype = 1)+
    # 8- Drawing the CP coordinates for Pab
  geom_point(data=D,aes(x = rpab_x, y = rpab_y),
             color= CP_colour,
             shape = 13,
             size = 3)+
    # 9-Drawing labels for each pab
  geom_text(data=D,aes(x = rpab_x, y = rpab_y),
                  color=CP_colour, 
                  label=str_c("P",{D$a_player_shirt},",",{D$b_player_shirt}),
                  max.overlaps=10,
                  size=3)+
    # 10- Drawing velocities of Pab
  geom_segment(data=D,aes(x = rpab_x, y = rpab_y, xend=rpab_x+vpab_x*s_v, yend=rpab_y+vpab_y*s_v),
              color=v_colour,
              arrow=arrow(length = unit(0.2,"cm")),
              lineend="round", linejoin = "round",
              linewidth=0.75)+
    # 11- Drawing accelerations of Pab
  geom_segment(data=D,aes(x = rpab_x, y = rpab_y, xend=rpab_x+apab_x*s_a, yend=rpab_y+apab_y*s_a),
                 color=a_colour,
                 arrow=arrow(length = unit(0.2,"cm")),
                 lineend="round", linejoin = "round")+
    # 12- Drawing the equivalent circle for Pab
  geom_circle(data=D,aes(x0=rpab_x, y0=rpab_y, r=sqrt(pab*s_P/pi)),
              inherit.aes = FALSE,
              n=25, alpha=0.1, color=CP_colour, fill=p_colour, linetype=2)+
    # 13-(W) Drawing Work of Pab
  geom_label(data=D,aes(x = rpab_x, y = rpab_y,fill=Wab),alpha=0.7,show.legend = FALSE,
                  color="black", 
                  #label=str_c("W",{D$a_player_shirt},",",{D$b_player_shirt},":",{round(D$Wab,2)}),
                  label={round(D$Wab,2)},
                  max.overlaps=10,
                  size=3)+
    # 14- Drawing the system variables (but take note that they aren't the CP variables)
  geom_label(data=DAB,aes(x = 0, y = -68/2-4),
                  color="black", 
                  label=str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2)),
                  size=6,
                  alpha=0.2)+
  
  ########### CENTRE OF PRESSURE ############
  
    # 15- Drawing the CP coordinates for PAB
  geom_point(data=D_CP_AB,aes(x = CP_rpAB_x, y = CP_rpAB_y),
             color= CP_colour,
             shape = 13,
             size = 3)+
    # 16- Drawing the variables for p-AB
  geom_circle(data=D_CP_AB,aes(x0=CP_rpAB_x, y0=CP_rpAB_y, r=sqrt(pAB*s_P/pi)),
              inherit.aes = FALSE,
              n=25, alpha=0.1, color=CP_colour, fill=p_colour, linetype=2, size=1)+
    # 17-Drawing labels for pAB
  geom_text(data=D_CP_AB,aes(x = CP_rpAB_x+1.5, y = CP_rpAB_y+1.5),
                  color=CP_colour,
                  label=str_c("pAB"),
                  max.overlaps=10,
                  size=4)+
    # 18- Drawing velocities of CP-Pab
  geom_segment(data=D_CP_AB,aes(x = CP_rpAB_x, y = CP_rpAB_y, xend=CP_rpAB_x+CP_vpAB_x*s_v, yend=CP_rpAB_y+CP_vpAB_y*s_v),
              color=v_colour,
              arrow=arrow(length = unit(0.2,"cm")),
              lineend="round", linejoin = "round",
              linewidth=0.75)+
  

    #Player size symbols 
  scale_size_manual(values=c(4, 4, 2))+
    #Player color for teams
  scale_color_manual(values=c("lightblue", "orange", "magenta"),aesthetics = "color", breaks = waiver())+
    #Work fill scale
  scale_fill_distiller(name = waiver(), aesthetics = "fill", palette = "Reds",direction = 1)+
    #Changingtheme options
  # guides(
  #   colour = guide_legend(position = "bottom"),
  #   size   = guide_legend(position = "bottom"),
  # ) +
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40)),
    #plot.subtitle = element_text(color = "black",size=12,
    #                          margin = margin(l = 50, b = -40))
    #plot.title = element_text(vjust=-0.5)
    #plot.caption = element_text(vjust = 0.5)
    #legend.margin = margin(0, 0, 0, 0), # turned off for alignment
    #legend.text = element_text(margin = margin(b = 0)),
    #legend.spacing.y = unit(0, "pt")
      #legend.position = 'none',
      #legend.location = "bottom",
            )+
  labs(
    title = str_c("frame: ",K[1,])
    #subtitle = str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

Play1

  #Creating transitions by layers
Play1_anim <- Play1 + 
  transition_layers( 
    layer_length = 2,
    transition_length = 0,
    keep_layers = TRUE,  #Permanecen todas las capas. Ejemplo: keep_layers = c(Inf, 0, 0) 
    from_blank = FALSE,  #No permanece la capa 1
    layer_order = NULL, #c(1,2,3,4,5,6,7,8,9,10,11,12),  #NULL si no queremos poner nada
    layer_names = NULL
    #layer_order = c(1,2,3,4,5,6,7,8,9,10,11,12)
  )
#enter_fade()  +  enter_grow()  +  exit_fade()  +  exit_shrink()
Play1_anim

# Render animation
Play1_video <- gganimate::animate(Play1_anim,
                       renderer = gifski_renderer(file = str_c(pathvideos,"Play1.gif")),
                       width = 1050, 
                       height = 680, 
                       nframes = 625,
                       fps = 25, 
                       duration = 25)
Play1_video
anim_save(str_c(pathvideos,"Play1.mp4"))

```

## Didactical. All frames

# HEATMAPS

## PAB+PBA

### PAB1

```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph111-WbA+WBa_Heatmap

  # (A1) counter-attack - FIRST Frame
# fr=31
# E <- EVENT_lost2 %>% filter(Id==fr)
# K <- KINEM %>% filter(Fr==E$Fr)
# D <- DYNAM %>% filter(Fr==E$Fr)
  # Selecting the frame to drawing

  #(A1) counter-attack - ALL Frames
E <- EVENT_lost2 %>% filter(Id==86)
K <- KINEM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)
D <- DYNAM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end & !is.nan(Wab))
DAB <- DYNAM %>% filter(Fr>=E$Fr & Fr<=E$Fr_end & !is.nan(Wab)) %>% group_by(Bowing) %>% 
  summarise(pAB=sum(pab), LAB=sum(Lab,na.rm=TRUE), WAB=sum(Wab,na.rm=TRUE)) %>% ungroup()

PlayHeatMap111 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab))+
    # 3- Drawing the headmap no-weighted
  #geom_density_2d_filled(aes(fill = after_stat(level)),alpha=0.5)+
  geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  geom_density_2d(aes(color = Wab)+
#   geom_density_2d_filled(
#     aes(group=a_player_shirt),
#     mapping = NULL,
#     data = NULL,
#     stat = "density_2d_filled",
#     position = "identity",
#     contour_var = "density",
#     na.rm = TRUE, #remove nan values
#     show.legend = NA,
#     inherit.aes = TRUE,
#     alpha=0.5
# )+
    #geom_density_2d(colour= "black",show.legend = FALSE)+

  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40)),
    #plot.subtitle = element_text(color = "black",size=12,
    #                          margin = margin(l = 50, b = -40))
    #plot.title = element_text(vjust=-0.5)
    #plot.caption = element_text(vjust = 0.5)
    #legend.margin = margin(0, 0, 0, 0), # turned off for alignment
    #legend.text = element_text(margin = margin(b = 0)),
    #legend.spacing.y = unit(0, "pt")
      #legend.position = 'none',
      #legend.location = "bottom",
            )+
  #Work fill scale
  #scale_fill_brewer(pallete="Reds", direction = 1)+
  #scale_fill_gradient(low = "white", high = "red", na.value = NA)+
  #scale_colour_distiller(pallete="Reds", direction = 1)+
  labs(
    title = str_c("frame: ",D$Fr[1]," to ", D$Fr[nrow(D)])
    #subtitle = str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap111

```

### PAB2

```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph112-WbA+WBa_Heatmap

  #counter-attack
E <- EVENT_lost2 %>% filter(Id==86)
K <- KINEM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end)
D <- DYNAM %>% filter (Fr>=E$Fr & Fr<=E$Fr_end & !is.nan(Wab))
DAB <- DYNAM %>% filter(Fr>=E$Fr & Fr<=E$Fr_end & !is.nan(Wab)) %>% group_by(Bowing) %>% 
  summarise(pAB=sum(pab), LAB=sum(Lab,na.rm=TRUE), WAB=sum(Wab,na.rm=TRUE)) %>% ungroup()
D <- D %>% select(Fr,a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap112 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  #annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab),alpha=0.3)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40))
    )+
  scale_colour_distiller(palette = "Reds")+
  labs(
    title = str_c("frame: ",D$Fr[1]," to ", D$Fr[nrow(D)])
    #subtitle = str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap112

```

### PAB3

```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph113-WbA+WBa_Heatmap

  #Analizing player 18(goalkeeper) and 11 Wab interaction
K <- KINEM %>% filter ((player_shirt==11 | player_shirt==18))
D <- DYNAM %>% filter (a_player_shirt==11 & b_player_shirt==18 & !is.nan(Wab))
DAB <- DYNAM %>% filter(a_player_shirt==11 & b_player_shirt==18 & !is.nan(Wab)) %>% group_by(Bowing) %>% 
  summarise(pAB=sum(pab), LAB=sum(Lab,na.rm=TRUE), WAB=sum(Wab,na.rm=TRUE)) %>% ungroup()
D <- D %>% select(Fr,a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap113 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab),alpha=0.7)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40))
    )+
  scale_colour_distiller(palette = "Reds",direction=1)+
  labs(
    title = str_c("frame: ",Fr_1I," to ", Fr_2F, ".  W18,11")
    #subtitle = str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap113

```

### PAB4

```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph114-WaBHeatmap

  #Analizing player 11 and team B Wab interaction
K <- KINEM %>% filter ((player_shirt==11))
D <- DYNAM %>% filter (a_player_shirt==11 & !is.nan(Wab) & Bowing=="Team B")
DAB <- DYNAM %>% filter(a_player_shirt==11 & !is.nan(Wab) & Bowing=="Team B") %>% group_by(Bowing) %>% 
  summarise(pAB=sum(pab), LAB=sum(Lab,na.rm=TRUE), WAB=sum(Wab,na.rm=TRUE)) %>% ungroup()
D <- D %>% select(Fr,a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap114 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab),alpha=0.7)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40))
    )+
  scale_colour_distiller(palette = "Reds",direction=1)+
  labs(
    title = str_c("frame: ",Fr_1I," to ", Fr_2F, "  W11,B")
    #subtitle = str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap114

```

### PAB5

```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph115-WaBHeatmap

  #Analizing player 11 and team B Wab interaction
K <- KINEM %>% filter ()
D <- DYNAM %>% filter (!is.nan(Wab) & Bowing=="Team B")
DAB <- DYNAM %>% filter(!is.nan(Wab) & Bowing=="Team B") %>% group_by(Bowing) %>% 
  summarise(pAB=sum(pab), LAB=sum(Lab,na.rm=TRUE), WAB=sum(Wab,na.rm=TRUE)) %>% ungroup()
D <- D %>% select(Fr,a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap115 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab),alpha=0.7)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40))
    )+
  scale_colour_distiller(palette = "Reds",direction=1)+
  labs(
    title = str_c("frame: ",Fr_1I," to ", Fr_2F, "  WA,B")
    #subtitle = str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap115

```

### PAB6

```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph116-paBHeatmap

  #Analizing player 11 and team B Wab interaction
K <- KINEM %>% filter ()
D <- DYNAM %>% filter (!is.nan(pab) & Bowing=="Team B")
DAB <- DYNAM %>% filter(!is.nan(pab) & Bowing=="Team B") %>% group_by(Bowing) %>% 
  summarise(pAB=sum(pab), LAB=sum(Lab,na.rm=TRUE), WAB=sum(Wab,na.rm=TRUE)) %>% ungroup()
D <- D %>% select(Fr,a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap116 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= pab),alpha=0.3)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40))
    )+
  scale_colour_distiller(palette = "Greens",direction=1)+
  labs(
    title = str_c("frame: ",Fr_1I," to ", Fr_2F, ".  pA,B")
    #subtitle = str_c("P= ",round({DAB$pAB},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap116

```


## WORK
### High defensive work. Team A
```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph120-Wab_HighHeatmap



%HighWork <- DYNAM %>% select(Fr,Bowing,rpab_x,rpab_y,pab,Wab,)
HighWork <- DYNAM %>% filter(!is.na(Wab)) %>% select(Wab) #472453 entradas
HighWork_percentil <-quantile(HighWork$Wab, probs = c(0.25, 0.50, 0.75, 0.90, 0.95, 0.99, 0.999), type = 7)
HighWork_percentil <- as.data.frame(HighWork_percentil)
#HighWork_percentil[5,1] # 95 percentil -> 0.2669686, 23623 entradas
#HighWork_percentil[6,1] # 99 percentil -> 0.6659084, 4725 entradas
#HighWork_percentil[7,1] # 99.9 percentil -> 5.4839910, 473 entradas

D <- DYNAM %>% filter(!is.na(Wab) & 
                        Wab >HighWork_percentil[6,1] &
                        Bowing=="Team B" &    #Note that team B has the possession of the ball
                        Wab < 50) %>% 
              select(Fr, a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap120 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab),alpha=0.7,show.legend = FALSE)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40))
    )+
  scale_colour_distiller(palette = "Reds",direction=1)+
  labs(
    title = str_c("High defensive work. Team A")
    #subtitle = str_c("P= ",round({D$Wab},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap120

```

### High defensive work. Team B
```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph121-Wab_HighHeatmap



#HighWork <- DYNAM %>% select(Fr,Bowing,rpab_x,rpab_y,pab,Wab,)
HighWork <- DYNAM %>% filter(!is.na(Wab)) %>% select(Wab) #472453 entradas
HighWork_percentil <-quantile(HighWork$Wab, probs = c(0.25, 0.50, 0.75, 0.90, 0.95, 0.99, 0.999), type = 7)
HighWork_percentil <- as.data.frame(HighWork_percentil)
#HighWork_percentil[5,1] # 95 percentil -> 0.2669686, 23623 entradas
#HighWork_percentil[6,1] # 99 percentil -> 0.6659084, 4725 entradas
#HighWork_percentil[7,1] # 99.9 percentil -> 5.4839910, 473 entradas

D <- DYNAM %>% filter(!is.na(Wab) & 
                        Wab >HighWork_percentil[6,1] &
                        Bowing=="Team A" &
                        Wab < 50) %>% 
              select(Fr, a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap121 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab),alpha=0.7,show.legend = FALSE)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40)),
    legend.position="bottom"
    )+
  scale_colour_distiller(palette = "Reds",direction=1)+
  labs(
    title = str_c("High defensive work. Team B")
    #subtitle = str_c("P= ",round({D$Wab},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap121

```

### High defensive work. Defending Player A6 (Attacking right midfielder)
```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph122-Wab_HighHeatmap



#HighWork <- DYNAM %>% select(Fr,Bowing,rpab_x,rpab_y,pab,Wab,)
HighWork <- DYNAM %>% filter(!is.na(Wab)) %>% select(Wab) #472453 entradas
HighWork_percentil <-quantile(HighWork$Wab, probs = c(0.25, 0.50, 0.75, 0.90, 0.95, 0.99, 0.999), type = 7)
HighWork_percentil <- as.data.frame(HighWork_percentil)
#HighWork_percentil[5,1] # 95 percentil -> 0.2669686, 23623 entradas
#HighWork_percentil[6,1] # 99 percentil -> 0.6659084, 4725 entradas
#HighWork_percentil[7,1] # 99.9 percentil -> 5.4839910, 473 entradas

D <- DYNAM %>% filter(!is.na(Wab) & 
                        Wab >HighWork_percentil[6,1] &
                        Bowing=="Team A" &
                        a_player_shirt == 6 &
                        Wab < 50) %>% 
              select(Fr, a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap122 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(color= Wab),alpha=0.9,show.legend = FALSE)+
  geom_text_repel(aes(label=paste("W", b_player_shirt, "," , a_player_shirt, sep = "")),alpha=0.4,size=1.75,show.legend = FALSE)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40)),
    legend.position="bottom"
    )+
  scale_colour_distiller(palette = "Reds",direction=1)+
  labs(
    title = str_c("Defending player 6.")
    #subtitle = str_c("P= ",round({D$Wab},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap122

```

### High defensive work. Player B30(Right Centre Midfielder) & Player A16(Attacking Centre Midfielder)
```{r}
#| echo: true
#| code-fold: true
#| label: fig-graph123-Wab_HighHeatmap

#HighWork <- DYNAM %>% select(Fr,Bowing,rpab_x,rpab_y,pab,Wab,)
HighWork <- DYNAM %>% filter(!is.na(Wab)) %>% select(Wab) #472453 entradas
HighWork_percentil <-quantile(HighWork$Wab, probs = c(0.25, 0.50, 0.75, 0.90, 0.95, 0.99, 0.999), type = 7)
HighWork_percentil <- as.data.frame(HighWork_percentil)
#HighWork_percentil[3,1] # 75 percentil -> 0.057008437, 29528,3125 entradas
#HighWork_percentil[5,1] # 95 percentil -> 0.2669686, 23623 entradas
#HighWork_percentil[6,1] # 99 percentil -> 0.6659084, 4725 entradas
#HighWork_percentil[7,1] # 99.9 percentil -> 5.4839910, 473 entradas

D <- DYNAM %>% filter(!is.na(Wab) & 
                        Wab >HighWork_percentil[3,1] & #Utilizamos el percentil 0.75 para representar 2 jugadores y que se puedan identificar la actividad defensiva 
                        #Bowing=="Team A" &
                        a_player_shirt == 16 &
                        b_player_shirt == 30 &
                        Wab < 50) %>% 
              select(Fr, Bowing, a_player_shirt,b_player_shirt,rpab_x,rpab_y,pab,Wab)

PlayHeatMap123 <- ggplot(D,aes(x=rpab_x,y=rpab_y)) +
    # 1-Drawing the pitch
  annotate_pitch(dimensions = Pitch_custom) +
    # 2-Drawing the CP
  geom_point(aes(size= Wab, color=Bowing),alpha=0.7,show.legend = FALSE)+
  #geom_density_2d(aes(size = Wab),alpha=0.3)+
  #geom_density_2d_filled(contour_var = "ndensity",alpha=0.5)+
  
  transparent_theme+
  theme(
    plot.title = element_text(color = "gray",size=12,
                              margin = margin(l = 50, b = -40)),
    legend.position="bottom"
    )+
  #scale_colour_distiller(palette = "Reds",direction=1)+
  labs(
    title = str_c("player 30 & player 16.")
    #subtitle = str_c("P= ",round({D$Wab},2),"  LT= ",round({DAB$LAB},2),"  W= ",round({DAB$WAB},2))
  )

PlayHeatMap123

```

# SIMULATING ACTIONS

Global variables for the model simulator

#REFERENCES
